% PHD THESIS ACKNOWLEDGEMENTS
% Michael Malick
% 2016-12-16


\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I sincerely thank my supervisors Randall Peterman and Sean Cox for their
guidance, advice, and encouragement. My success reflects their ability as
mentors and I am grateful for their willingness to allow me the freedom to
explore new paths. To Randall, you provided an exceptional example of how to be
a great scientist and I will always be grateful for the opportunities you
provided me at SFU. To Sean, you have been a source of inspiration as a
scientist and I am grateful for you welcoming me into your lab.

I have benefited greatly from the advice and many insightful discussions with my
supervisory committee. I am grateful to Franz Mueter for being available to
discuss my research and taking the time to provide critical feedback. I thank
Murray Rutherford for his patience and guidance as I learned the details of
policy research and his clear and reliable insights into the policy process.

I am grateful to the individuals who participated as collaborators in my
research. Thanks to Thomas Wainwright and Bill Peterson, who hosted me in
Newport, OR, and provided keen insights into the Oregon Coast marine ecosystem.
I also thank Nathan Mantua and Todd Mitchell who hosted me at the University of
Washington and taught me the details of working with satellite data sets.

To my fellow lab members in the fisheries group at SFU, thank you for providing
me a source of companionship and your willingness and enthusiasm to discuss
anything quantitative. I am grateful to my colleagues in the REM PhD program who
provided inspiration and a sounding board for me to discuss my research.

I thank the many biologists from the Washington Department of Fish and Wildlife,
Fisheries and Ocean Canada, the Alaska Department of Fish and Game, and the
Newport Research Station for collecting and providing me with the numerous data
sets analyzed here. Without their efforts, this research would not have been
possible.

I am grateful for funding from the Canada Research Chairs Program provided to
Randall Peterman and for the numerous scholarships and fellowships provided by
Simon Fraser University that helped support my research.

It is beyond my abilities to thank my family for their encouragement during my
education. To my beautiful wife Erin, thank you for providing me a source of joy
and happiness everyday and your willingness to listen to me discuss the details
of my research. To my parents, brother, and extended family, thank you for your
unwavering support through all my years of education. Mother and Dad, you have
always given me the freedom to pursue my interests and provided me with
unconditional encouragement, thank you.

My life has be greatly enriched by these individuals, and for that I am
grateful.


