% LME Coefficient Table
% 2016-10-31

{\small
  \libertineLF

  \begin{ThreePartTable}
    \begin{TableNotes}
      {\footnotesize
        \item[*] Significantly different from zero at P \textless{} 0.05. 
        \item[**] Significantly different from zero at P \textless{} 0.01. 
      }
    \end{TableNotes}

    \begin{longtable}{llllllllll}
      \caption[Summary of the best-fit multi-stock Ricker model
        coefficients.]{Summary of the best-fit multi-stock Ricker model
        coefficients (eq. \ref{eq:bloom:3}). The Subset column identifies the
        stocks included in the hierarchical model (North = stocks 1-9, South =
        stocks 10-27, and All = stocks 1-27). The Stocks column indicates the
        number of stocks used to fit the model; N is the total number of data
        points across all stocks used to fit the model; $\alpha$ is the
        intercept representing average productivity of pink salmon at low
        spawner abundance (fixed effect; see eq. \ref{eq:bloom:3}); $\gamma$ is
        the fixed effect corresponding to either the bloom initiation or mean
        chl-a concentration (see eq. \ref{eq:bloom:3}); and SE is the standard
        error for the fixed-effect coefficients. AIC\textsubscript{C} is the
        Akaike Information Criterion, corrected for small sample size.} \\ 
      \hline
      \input{./3_springbloom/tables/R_table_lme.tex}
      \hline
      \insertTableNotes
      \label{tab:bloom:2}
    \end{longtable}

  \end{ThreePartTable}
}

% \noindent * Significantly different from zero at P \textless{} 0.05; **
% significant at P \textless{} 0.01.
