% PHD THESIS ACKNOWLEDGEMENTS
% Michael Malick
% 28 August 2015


\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I sincerely thank my supervisors Randall Peterman and Sean Cox for their
guidance, advice, and encouragement. My success reflects their ability as
mentors and I am grateful for their inspiration and their willingness to allow
me the freedom to explore new paths. To Randall, you provided an exceptional
example of how to be a great scientist and I will always be grateful for the
opportunities you provided me at SFU. To Sean, you have been a source of
inspiration as a scientist and I am grateful for you welcoming me into your lab.

I have benefited greatly from the advice and  many insightful discussion with my
supervisory committee. I am grateful to Franz Mueter for being available to
discuss my research and taking the time to provide critical feedback on my
research.  I thank Murray Rutherford for his patience and guidance as I learned
the details of policy research and his clear and reliable insights to the policy
process.

I am grateful to the many people who participated as collaborators in my
research.  Thanks to Thomas Wainwright and Bill Peterson, who hosted me in
Newport, OR, and provided keen insights and guidance on the marine system off
the coast of Newport. I also thank Nathan Mantua and Todd Mitchell who hosted me
at the University of Washington and taught me the details of working with
satellite data sets.

To my fellow lab member in the fisheries group at SFU, thank you for providing
me a source of companionship and your willingness and enthusiasm to discuss
anything quantitative. I am grateful to my colleagues in the REM PhD program who
provided inspiration and a sounding board for me to discuss my research.

I thank the many biologists from the Washington Department of Fish and Wildlife,
Fisheries and Ocean Canada, the Alaska Department of Fish and Game, and the
Newport Research Station for collecting and providing me with the numerous data
sets analyzed here.  Without their efforts this research would not have been
possible.  

I am grateful for funding from the Canada Research Chairs Program provided to
Randall Peterman and to Simon Fraser University for the numerous scholarships
and fellowships that also supported my research.

It is beyond my abilities to thank my family for their encouragement during my
education. To my beautiful wife Erin, thank you for providing me a source of joy
and happiness everyday and your willingness to listen to me discuss the details
of my research. To my parents, brother, and extended family, thank you for your
unwavering support through all my years of education.  Mother and Dad, you have
always given me the freedom to pursue my interests and provided me with
unconditional encouragement, thank you.

My life has be greatly enriched by these individuals, and for that I am
grateful.


